---
title: "02_street_segments"
author: "Iain Agar"
date: "11/05/2021"
output: html_document
---

# Create a street segments map using features from OSM

OSM features available for geospatial analysis can be viewed here https://wiki.openstreetmap.org/wiki/Map_features

We can see which features are available, the index url above for osm identifies that we can find road data under the categories of highway.

Results show the various features contained within highway. Of interest for our crime by street segments, or hot routes, we are most likely interested in:

* primary
* secondary
* tertiary
* residential
* track
* motorway
* trunk

```{r}

# see which sub-categories fall within the highway features
available_tags("highway")

```

## Creating a bounding box for your study area

For this example I am using the city of Baltimore, Maryland, US.

```{r}

bbox_1 <- getbb("Baltimore") %>%
  # get bounding box of Baltimore
  opq() %>%
  # add feature
  add_osm_feature("highway")

# check out the structure
str(bbox_1)

```

## Create a street segment simple feature

We want a polyline of the street network in Baltimore.

```{r}

# create a simple features object from bbox_1
b_md_roads <- osmdata_sf(bbox_1)

# see object classes, will show a simple features collection by number of points, lines, polygons etc
class(b_md_roads)

# we want the polylines
b_md_roadsPL <- b_md_roads$osm_lines

```

## Data cleaning

The osm datasets are very wide, but we are only interested in a small number of columns for this process:

* osm_id
* name
* highway
* geometry

```{r}

# subset data
br <- b_md_roadsPL[,c("osm_id", "name", "highway", "geometry")]

# add length of segments
br$length <- as.numeric(st_length(br$geometry))

# subset the highway features of interest
baltimore_roads <- subset(br, highway=="motorway" | highway=="primary" | highway=="residential" | highway=="secondary" | highway=="tertiary" | highway=="track" | highway=="trunk")
```